Home Nighttime Cough Monitoring

### 2025-02-05 Lilah Melzer/cough_monitor_analysis.qmd
need to add CFQ-R plots

```{r setup}
#| include: false
library(Hmisc)
library(redcapAPI)
library(rspiro)
library(childsds)
library(arsenal)
library(lmerTest)
library(performance)
library(GLMMadaptive)
library(tidyverse)
library(ggstats)
```

@@

start_end <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/Start and end dates.csv",

@@

vacations <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/Vacation days.csv",

@@

  "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Raw/Cough_w_TotalCounts_1105_NMDcorrected - Cough_agg_w_count_122324.csv",

@@

    sid, Diagnosis, age_at_consent, height, weight, pre_fev,

@@

  select(sid, age_at_consent, height, weight, nih_sex, pre_fev) %>%

@@

@@

  left_join(demo, fev1[, c("sid", "height_perc", "weight_perc", "pre_fev_pp")])

@@

@@

@@

  select(sid, height_perc, weight_perc, pre_fev_pp, everything()) %>%

@@

  file = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor/Data_Clean/cleaned_aggregated_counts.csv",

@@

@@

cfqr_6_13 <- df %>%
  filter(redcap_repeat_instrument == "CFQ-R Caregivers 6-13") %>%
  select(
    sid, Diagnosis, cfqr_caregivers_613_timestamp:cfqr_caregivers_613_complete
  )

cfqr_6_13$cfqr_physical <- apply(
  data.matrix(cfqr_6_13[, paste0("cfq_r_p_", 1:5)]), 1, function(r) {
    (mean(r) - 1) / 3 * 100
  }
)
cfqr_6_13$cfqr_emotion <- apply(
  data.matrix(cfqr_6_13[, paste0("cfq_r_p_", 6:13)]), 1, function(r) {
    (mean(r) - 1) / 3 * 100
  }
)

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(redcapAPI)
library(rspiro)
library(childsds)
library(arsenal)
library(lmerTest)
library(performance)
library(GLMMadaptive)
library(tidyverse)
library(ggstats)
library(corrplot)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor")
```

@@

start_end <- read.csv("./Data_Clean/Start and end dates.csv",

@@

vacations <- read.csv("./Data_Clean/Vacation days.csv",

@@

  "./Data_Raw/Cough_w_TotalCounts_1105_NMDcorrected - Cough_agg_w_count_122324.csv",

@@

    sid, Diagnosis, age_at_consent, height, weight, pre_fev, post_fev,

@@

  select(sid, age_at_consent, height, weight, nih_sex, pre_fev, post_fev) %>%

@@

fev1$post_fev_pp <- pctpred_GLIgl(fev1$age_at_consent, fev1$height / 100,
  FEV1 = fev1$post_fev

@@

  left_join(
    demo,
    fev1[, c("sid", "height_perc", "weight_perc", "pre_fev_pp", "post_fev_pp")]

@@

label(demo$pre_fev_pp) <- "Post-bronchodilator FEV1 % Predicted"

@@

# Import and merge sleep times from Lilah
sleep <- read.csv("./Data_Clean/Total sleep times.csv", na.strings = "")
sleep <- sleep %>%
    Participant.First.Name..or.any.uniq.identifier.,
    Total.sleep.time..hours.
  rename(
    sid = Participant.First.Name..or.any.uniq.identifier.,
    total_sleep = Total.sleep.time..hours.
sleep$sid <- as.character(as.numeric(sub("FN", "", sleep$sid)))
label(sleep$total_sleep) <- "Total sleep time (hours)"
demo <- left_join(demo, sleep)

@@

  select(sid, height_perc, weight_perc, pre_fev_pp, total_sleep, everything()) %>%

@@

  file = "./Data_Clean/cleaned_aggregated_counts.csv",

@@

# Basic statistics excluding 0s

## Aggregated by person

The comparisons below exclude any nights with missing data OR 0 coughs, as well as periods of sickness. This table could potentially be a little bit confusing because the numbers presented are the median of multiple medians, but the data are still so skewed without 0s that this seemed the best way to aggregate the data.

```{r results='asis'}
#| warning: false
no_0 <- agg_counts %>%
  filter(Sick == "No", !is.na(Coughs)) %>%
  group_by(sid, Diagnosis) %>%
  mutate(`Prop. Non-0 Coughs` = mean(Coughs > 0, na.rm = T)) %>%
  filter(Coughs > 0) %>%
  summarise(
    `Prop. Non-0 Coughs` = `Prop. Non-0 Coughs`[1],
    `Median Coughs` = median(Coughs),
    `Cough IQR` = IQR(Coughs),
    .groups = "drop"
  )
no_0 <- left_join(
  no_0,
  fev1 %>%
    mutate(sid = as.numeric(sid)) %>%
    select(sid, pre_fev_pp, post_fev_pp),
  by = join_by(sid)
)
no_0_t <- tableby(
  Diagnosis ~
    kwt(`Prop. Non-0 Coughs`, "medianq1q3", "range") +
    kwt(`Median Coughs`, "medianq1q3", "range"),
  data = no_0
)
summary(no_0_t, pfootnote = T)
M <- cor.test(no_0$`Median Coughs`, no_0$pre_fev_pp, method = "spearman")
corrplot(cor(no_0 %>% select(`Median Coughs`, pre_fev_pp),
  method = "spearman", use = "pairwise.complete.obs"
))
ggplot(no_0, aes(x = `Median Coughs`, y = pre_fev_pp)) +
  geom_point()
```

Spearman correlation between median coughs and pre-bronchodilator FEV1 was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,eps = 0.001,digits = 3)`).

## Not aggregated

```{r results='asis'}
#| warning: false
no_0 <- agg_counts %>% filter(Sick == "No", !is.na(Coughs), Coughs > 0)
no_0 <- left_join(
  no_0,
  fev1 %>%
    mutate(sid = as.numeric(sid)) %>%
    select(sid, pre_fev_pp, post_fev_pp),
  by = join_by(sid)
)
no_0_t <- tableby(Diagnosis ~ kwt(Coughs, "medianq1q3", "range"), data = no_0)
summary(no_0_t, pfootnote = T)
```

Spearman correlation between median coughs and pre-bronchodilator FEV1 was `r round(as.numeric(M$estimate),3)` (p = `r format.pval(M$p.value,eps = 0.001,digits = 3)`).

@@

# Caregivers of children ages 6-13
# Rename columns so we can use the code from the CFF
cfqr_parent_6_13_names <- list(
  "cfq_r_p_1" = "phys1", "cfq_r_p_2" = "phys2", "cfq_r_p_3" = "phys3",
  "cfq_r_p_4" = "phys4", "cfq_r_p_5" = "phys5", "cfq_r_p_6" = "emot6",
  "cfq_r_p_7" = "emot7", "cfq_r_p_8" = "vital8", "cfq_r_p_9" = "vital9",
  "cfq_r_p_10" = "vital10", "cfq_r_p_11" = "vital11", "cfq_r_p_12" = "vital12",
  "cfq_r_p_13" = "school13", "cfq_r_p_14" = "phys14", "cfq_r_p_15" = "phys15",
  "cfq_r_p_16" = "phys16", "cfq_r_p_17" = "eat17", "cfq_r_p_18" = "treat18",
  "cfq_r_p_19" = "body19", "cfq_r_p_20" = "body20", "cfq_r_p_21" = "body21",
  "cfq_r_p_22" = "health22", "cfq_r_p_23" = "emot23", "cfq_r_p_24" = "health24",
  "cfq_r_p_25" = "emot25", "cfq_r_p_26" = "emot26", "cfq_r_p_27" = "school27",
  "cfq_r_p_28" = "school28", "cfq_r_p_29" = "school29",
  "cfq_r_p_30" = "treat30", "cfq_r_p_31" = "treat31", "cfq_r_p_32" = "health32",
  "cfq_r_p_33" = "weight33", "cfq_r_p_34" = "resp34", "cfq_r_p_35" = "resp35",
  "cfq_r_p_36" = "resp36", "cfq_r_p_37" = "resp37", "cfq_r_p_38" = "resp38",
  "cfq_r_p_39" = "resp39", "cfq_r_p_40" = "resp40", "cfq_r_p_41" = "digest41",
  "cfq_r_p_42" = "digest42", "cfq_r_p_43" = "digest43", "cfq_r_p_44" = "eat44"
cfqr_parent_6_13_names <-
  setNames(names(cfqr_parent_6_13_names), cfqr_parent_6_13_names)
# Some items need to be reverse scored
cfqr_parent_6_13_reverse <- c(
  "emot6", "vital10", "vital12", "phys15", "treat31", "health22", "health24",
  "health32", "school28", "resp37"
# Do it!
cfqr_parent_6_13 <- df %>%
    sid, redcap_event_name:redcap_repeat_instance,
    cfq_r_p_1:cfq_r_p_44
  mutate(across(cfq_r_p_1:cfq_r_p_44, ~ as.numeric(.x))) %>%
  rename(all_of(unlist(cfqr_parent_6_13_names))) %>%
  mutate(across(all_of(cfqr_parent_6_13_reverse), ~ 5 - .x)) %>%
    physical =
      ifelse(
        sum(is.na(c(phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16))) <= 4,
        (mean(c(phys1, phys2, phys3, phys4, phys5, phys14, phys15, phys16)) - 1)
        / 3 * 100, NA
    emotion =
      ifelse(
        sum(is.na(c(emot6, emot7, emot23, emot25, emot26))) <= 2,
        (mean(c(emot6, emot7, emot23, emot25, emot26)) - 1)
        / 3 * 100, NA
    vitality =
      ifelse(
        sum(is.na(c(vital8, vital9, vital10, vital11, vital12))) <= 2,
        (mean(c(vital8, vital9, vital10, vital11, vital12)) - 1)
        / 3 * 100, NA
    school =
      ifelse(
        sum(is.na(c(school13, school27, school28, school29))) <= 2,
        (mean(c(school13, school27, school28, school29)) - 1)
        / 3 * 100, NA
    eat =
      ifelse(
        sum(is.na(c(eat17, eat44))) == 0,
        (mean(c(eat17, eat44)) - 1)
        / 3 * 100, NA
    body =
      ifelse(
        sum(is.na(c(body19, body20, body21))) <= 1,
        (mean(c(body19, body20, body21)) - 1)
        / 3 * 100, NA
    treat =
      ifelse(
        sum(is.na(c(treat18, treat30, treat31))) <= 1,
        (mean(c(treat18, treat30, treat31)) - 1)
        / 3 * 100, NA
    health =
      ifelse(
        sum(is.na(c(health22, health24, health32))) <= 1,
        (mean(c(health22, health24, health32)) - 1)
        / 3 * 100, NA
    respirat =
      ifelse(
        sum(is.na(c(resp34, resp35, resp36, resp38, resp39, resp40))) <= 3,
        (mean(c(resp34, resp35, resp36, resp38, resp39, resp40)) - 1)
        / 3 * 100, NA
    digest =
      ifelse(
        sum(is.na(c(digest41, digest42, digest43))) <= 1,
        (mean(c(digest41, digest42, digest43)) - 1)
        / 3 * 100, NA
    weight =
      ifelse(
        sum(is.na(c(weight33))) == 0,
        (mean(c(weight33)) - 1)
        / 3 * 100, NA
check <- left_join(
  df %>%
      sid, redcap_event_name:redcap_repeat_instance,
      cfq_r_p_1:cfq_r_p_44
  cfqr_parent_6_13 %>%
    select(sid, redcap_event_name:redcap_repeat_instance, physical:weight),
  join_by(
    sid, redcap_event_name, redcap_repeat_instrument,
    redcap_repeat_instance
write.csv(check, file = "./Data_Clean/cfqr_parent_6_13_check.csv", na = "", row.names = F)

cfqr_children_6_11 <- list(
  "cfq_r_y_1" = "phys1", "cfq_r_y_2" = "phys2", "cfq_r_y_3" = "phys3",
  "cfq_r_y_4" = "phys4", "cfq_r_y_5" = "phys5", "cfq_r_y_6" = "phys6",
  "cfq_r_y_7" = "emot7", "cfq_r_y_8" = "emot8", "cfq_r_y_9" = "emot9",
  "cfq_r_y_10" = "emot10", "cfq_r_y_11" = "emot11", "cfq_r_y_12" = "emot12",
  "cfq_r_y_13" = "emot13", "cfq_r_y_14" = "emot14", "cfq_r_y_15" = "eat15",
  "cfq_r_y_16" = "treat16", "cfq_r_y_17" = "eat17", "cfq_r_y_18" = "treat18",
  "cfq_r_y_19" = "eat19", "cfq_r_y_20" = "social20", "cfq_r_y_21" = "social21",
  "cfq_r_y_22" = "social22", "cfq_r_y_23" = "social23",
  "cfq_r_y_24" = "social24", "cfq_r_y_25" = "social25",
  "cfq_r_y_26" = "social26", "cfq_r_y_27" = "body27", "cfq_r_y_28" = "body28",
  "cfq_r_y_29" = "body29", "cfq_r_y_30" = "treat30", "cfq_r_y_31" = "resp31",
  "cfq_r_y_32" = "resp32", "cfq_r_y_33" = "resp33", "cfq_r_y_34" = "resp34",
  "cfq_r_y_35" = "digest35 "
)
cfqr_children_6_11 <-
  setNames(names(cfqr_children_6_11), cfqr_children_6_11)
cfqr_teens_adults <- list(
  "cfq_r_a_1" = "phys1", "cfq_r_a_2" = "phys2", "cfq_r_a_3" = "phys3",
  "cfq_r_a_4" = "phys4", "cfq_r_a_5" = "phys5", "cfq_r_a_6" = "vital6",
  "cfq_r_a_7" = "emot7", "cfq_r_a_8" = "emot8", "cfq_r_a_9" = "vital9",
  "cfq_r_a_10" = "vital10", "cfq_r_a_11" = "vital11", "cfq_r_a_12" = "emot12",
  "cfq_r_a_13" = "phys13", "cfq_r_a_14" = "eat14", "cfq_r_a_15" = "treat15",
  "cfq_r_a_16" = "treat16", "cfq_r_a_17" = "treat17", "cfq_r_a_18" = "health18",
  "cfq_r_a_19" = "phys19", "cfq_r_a_20" = "phys20", "cfq_r_a_21" = "eat21",
  "cfq_r_a_22" = "social22", "cfq_r_a_23" = "social23", "cfq_r_a_24" = "body24",
  "cfq_r_a_25" = "body25", "cfq_r_a_26" = "body26", "cfq_r_a_27" = "social27",
  "cfq_r_a_28" = "social28", "cfq_r_a_29" = "social29",
  "cfq_r_a_30" = "social30", "cfq_r_a_31" = "emot31", "cfq_r_a_32" = "health32",
  "cfq_r_a_33" = "emot33", "cfq_r_a_34" = "health34", "cfq_r_a_35" = "role35",
  "cfq_r_a_36" = "role36", "cfq_r_a_37" = "role37", "cfq_r_a_38" = "role38",
  "cfq_r_a_39" = "weight39", "cfq_r_a_40" = "resp40", "cfq_r_a_41" = "resp41",
  "cfq_r_a_42" = "resp42", "cfq_r_a_43" = "resp43", "cfq_r_a_44" = "resp44",
  "cfq_r_a_45" = "resp45", "cfq_r_a_46" = "resp46", "cfq_r_a_47" = "digest47",
  "cfq_r_a_48" = "digest48", "cfq_r_a_49" = "digest49", "cfq_r_a_50" = "eat50"
)
cfqr_teens_adults <-
  setNames(names(cfqr_teens_adults), cfqr_teens_adults)
```
